---
title: "Case Study one"
author: "Dusin Bracy and Tazeb Abera"
date: "10/8/2019"
output: html_document
---

# Case study 1(Case Study of US Craft Beer and Breweries)

## Introduction and Overview

 
-------

### Repository Structure

### Reproduction of Analysis

## Analysis

We begin by setting `knitr` options, 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r , include=FALSE}
library(ggplot2) # Data visualization
library(readr) # CSV file I/O, e.g. the read_csv function
library(dplyr) # Heavy use of this library
library(tidyr)
library(ggthemes) # for themes in ggplot
library(knitr)
library(kableExtra) # library to make the document more presentable
```

```{r}
# Read in beer and breweries data set using _csv for more tidy output
Beers <- read_csv('Beers.csv', col_types = cols())
Breweries <- read_csv('Breweries.csv', col_types = cols())
```

The field `Brewery_id` in `Beers.csv` and `Brew_ID` in `Breweries.csv` are the same data, but do not share a name. We remedy this by renaming the column in `Beers.csv`.

## Research Questions


### 1. How many breweries are present in each state?

We answer this question by retrieving the value of `State` from the `Breweries` data.
```{r}
BrewPerState <-table(Breweries$State)
BrewPerState
```


```{r}
# Renamed Brewery_id to Brew_ID to satisfy merging requirement
Beers <- rename(Beers, Brew_ID = Brewery_id)
```
### 2. Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.

 We can now merge them into a single dataset, using `full_join`.
 
```{r}
Brewdata <- full_join(Beers, Breweries, by="Brew_ID")
```

```{r}
# Change variable names to more meaningful title
Brewdata <- rename(Brewdata, Beer = Name.x, Brewery = Name.y, OZ = Ounces)
```

To retrieve the first and last six observations from the combined data, 
we run `head` and `tail` on `Brewdata`, our combined dataset.
```{r}
kable(Brewdata %>% head())
```

```{r}
kable(Brewdata %>% tail())
```